---
title: "coefficient of variation"
author: "Elizabeth Forbes"
date: "1/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

UPDATE: January 26, 2021. Kelly and I worked on the above problem in Python, using a Jupyter notebook, and output the data (e.g. pulling the closest fluxbot flux for each CIRAS measurement and pairing them) into the github repo for our project. The file is called "ciras_fluxbot_calval.csv".

Update Update: Oct. 8th, 2021.  I ran the notebook again just to confirm that the data for the fluxbots was the most updated version bc I couldn't be sure from my process notes.  I think it's the same.  Womp womp.  Using most updated version of the paired datasets here.

```{r calval_jan26}
library(tidyverse)

# import new dataset
calval_oct <- read_csv("ciras_fluxbot_calval_2.csv")
View(calval_oct)

# filter by those rows where delta_T is less than 25mins and qaqc is less than 11
library(dplyr)
calval_oct_b <- calval_oct %>% 
  filter(delta_T < "0 days 00:25:00") %>% 
  filter(qaqc_flags < 11) # 75 had >11 for qaqc flags, removed (137 total)
# remove those with OM as feature
calval_oct_c <- calval_oct_b %>% 
  filter(location == "Open Soil" | location == "Under Tree") # 86 total observations left

```


## Analytical comparison
After importing data, explore the data.
Questions to ask: 
- is there a consistent pattern in the relationship between CIRAS and corresponding fluxbot values depending on whether CIRAS was taken before or after a fluxbot value? (E.G. potential impact of data collector being present and breathing around the open fluxbot)
- is there a predictable relationship between dt and fidelity of CIRAS/fluxbot values?
- etc.

```{r data exploration}
library(tidyr)
library(ggridges)
library(ggplot2)

# simple ridge plot of each set of data: first, fluxbots
p1 <- ggplot(calval_oct_c, aes(x = flux_umol_m2_sec, y=BOTID, fill=TRTMNT))+
  geom_density_ridges(alpha=.25)+
  theme_ridges()
  # xlim(-6,10) # cutting off the extremes so i can look at the spread a little easier
  #theme(legend.position = "none")

# simple ridge plot of each set of data: second, CIRAS linear fit
p2 <- ggplot(calval_oct_c, aes(x = CIRASFLX_L, y=BOTID, fill=TRTMNT))+
  geom_density_ridges(alpha=.25)+
  theme_ridges()
  #theme(legend.position = "none")

# simple ridge plot of each set of data: first, CIRAS quadratic fit
p3 <- ggplot(calval_oct_c, aes(x = CIRASFLX_Q, y=BOTID, fill=TRTMNT))+
  geom_density_ridges(alpha=.25)+
  theme_ridges()
  #theme(legend.position = "none")

# put them all together:
library(gridExtra)
grid.arrange(p1, p2, p3, ncol=1, nrow=3)

# of course, the x axes are all off: the spread for the fluxbots is higher.
```

Okay so the above just had us looking at the distribution.  Want to do the same scatterplots I did before with the old data, to see how well-aligned the two methods are visually.

```{r scatterplots}
# SCATTERPLOTS of the data: the next four graphs are just the fluxbot data against the CIRAS data.

# plot fluxbot data against CIRAS linear fit data, colored by treatment:
ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_L, color=TRTMNT))+
  geom_point(alpha=.75)+
  theme_classic()

# plot fluxbot data against CIRAS quadratic fit data, colored by treatment:
ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_Q, color=TRTMNT))+
  geom_point(alpha=.75)+
  theme_classic()

# plot fluxbot data against CIRAS linear fit data, colored by feature:
ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_L, color=FEAT))+
  geom_point(alpha=.75)+
  theme_classic()

# plot fluxbot data against CIRAS quadratic fit data, colored by feature:
ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_Q, color=FEAT))+
  geom_point(alpha=.75)+
  theme_classic()
```

More scatterplots, but side-by-side:
```{r scatterplots2}

# let's add SIZE based on the amount of time in minutes elapsed between CIRAS and fluxbot data:
g1 <- ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_L, color=TRTMNT))+
  geom_point(alpha=.35)+
  theme_classic()+
  scale_size_continuous(guide=FALSE)+
  labs(x="fluxbot soil flux",
       y="CIRAS soil flux, LINEAR regression")+
  geom_abline(intercept = 0, slope = 1, linetype=2, colour="red")

g2 <- ggplot(data=calval_oct_c, aes(x=flux_umol_m2_sec, y=CIRASFLX_Q, color=TRTMNT))+
  geom_point(alpha=.35)+
  theme_classic()+
  scale_size_continuous(guide=FALSE)+
  labs(x="fluxbot soil flux",
       y="CIRAS soil flux, QUADRATIC regression")+
  geom_abline(intercept = 0, slope = 1, linetype=2, colour="red")

library(ggpubr)
g3 <- ggarrange(g1, g2, ncol=2, nrow=1, common.legend=TRUE, legend = "bottom")

annotate_figure(g3,top = text_grob("Two methods for measuring soil CO2 flux", color = "coral", face = "bold", size = 14),
                bottom = text_grob("point size indicates minutes elapsed between measurements", color = "black", size = 10))

# just linear calc for CIRAS (they are similar in aggregate):
annotate_figure(g1,top = text_grob("Two methods for measuring soil CO2 flux", color = "coral", face = "bold", size = 14),
                bottom = text_grob("point size indicates minutes elapsed between measurements", color = "black", size = 10))
```

Look just by treatment:
```{r scatterplot3}

# look just by treatment:
library(dplyr)
library(ggplot2)
# subset data: first, MWC:
calval_oct_c %>% 
  filter(TRTMNT == "MWC") %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=CIRASFLX_L))+
  geom_point(alpha=.35, colour="coral1")+
  theme_classic()+
  facet_wrap(~BOTID)+
  scale_size_continuous(guide=FALSE)+
  labs(x="fluxbot soil flux",
       y="CIRAS soil flux, LINEAR regression")+
  theme(legend.position = "none")+
  #xlim(0,10)+
  ylim(0,8)+
  geom_abline(intercept = 0, slope = 1, linetype=2, colour="red")+
  ggtitle("NMWC fluxbots")

# second, O:
calval_oct_c %>% 
  filter(TRTMNT == "O") %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=CIRASFLX_L))+
  geom_point(alpha=.35, colour="cyan3")+
  theme_classic()+
  facet_wrap(~BOTID)+
  scale_size_continuous(guide=FALSE)+
  labs(x="fluxbot soil flux",
       y="CIRAS soil flux, LINEAR regression")+
  theme(legend.position = "none")+
  # xlim(0,40)+
  # ylim(0,8)
  geom_abline(intercept = 0, slope = 1, linetype=2, colour="red")+
  ggtitle("O fluxbots")
```

Calculate correlation coefficient:
```{r analysis}

# calcaulate corr coeff:
cor(calval_oct_c$flux_umol_m2_sec, calval_oct_c$CIRASFLX_L)
#[1] -0.2923191

# take a look at at this data (the calval data minus the MWC rep 1 bots):
ggplot(data=calval_jan26_c, aes(x=flux_umol_m2_sec, 
                                y=CIRASFLX_L, color=TRTMNT))+
  geom_point(alpha=.35)+
  theme_classic()+
  facet_wrap(~FEAT)+
  scale_size_continuous(guide=FALSE)+
  labs(x="fluxbot soil flux",
       y="CIRAS soil flux, LINEAR regression")+
  geom_abline(intercept = 0, slope = 1, linetype=2, colour="red")
  # xlim(0,10)+
  # ylim(0,8)

```

```{r descriptive_stats}
# calculate some descriptive statistics for inclusion in supplement.

#CIRAS data, linear and quadratic:
mean(calval_oct_c$CIRASFLX_L) # 1.583605
mean(calval_oct_c$CIRASFLX_Q) # 1.421279

# mean and median of the fluxbot data corresponding to the CIRAS data:
mean(calval_oct_c$flux_umol_m2_sec)# 3.535814
median(calval_oct_c$flux_umol_m2_sec) # 2.595

# median of linear CIRAS data
median(calval_oct_c$CIRASFLX_L) # 1.38

# coefficient of variation, or measure of dispersion of the data around the mean (ratio of standard deviation to the mean):
library(EnvStats)
cv(calval_oct_c$CIRASFLX_L) #0.4681408
cv(calval_oct_c$flux_umol_m2_sec) #0.9737557

```