---
title: "finalanalyses_fluxbot_Feb2021"
author: "Elizabeth Forbes"
date: "2/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(DHARMa)
```


## Analyses for fluxbot methods paper, using final dataset

This document contains the analyses we'll use for our fluxbots methods paper. Using the fluxdat_final dataset constructed in "finaldataset_fluxbots_Feb2021" for all analyses.

# basic exploration:
```{r basic}
mean(fluxdat_final$flux_umol_m2_sec)
# [1] 3.971561
## updated 09/15/2022: 3.10518
#### updated 04/18/2023: 4.050193
median(fluxdat_final$flux_umol_m2_sec)
# 3.144

min(fluxdat_final$flux_umol_m2_sec)
# -3.893
## updated 09/15/2022: -2.655
#### updated 04/18/2023:-3.695

max(fluxdat_final$flux_umol_m2_sec)
# 43.313
## updated 09/15/2022: 38.51
#### updated 04/18/2023: 46.395

sum(fluxdat_final$flux_umol_m2_sec < 0) # only 334 of dataset's final fluxes are negative
sum(fluxdat_final$flux_umol_m2_sec < -1) # only 42 of dataset's final 'negative' fluxes are less than -1

```


# comparisons, night vs. day data
This chunk has some brief comparisons between data collected at night, and during the day, for both flux and ambient CO2 values.
```{r ttests}
# View(fluxdat_final)

# want to first compare all midnight data to all high noon data. In all cases, I want to be able to put down bot location as a random effect; so this is going to be more of linear mixed effects model than a straight t-test.
# run a paired (e.g. not independent) two-sample t-test on these data:
midnight <- subset(fluxdat_final, hour == 0) # subset midnight data
noon <- subset(fluxdat_final, hour == 12) # subset noon data

t1 <- t.test(midnight$flux_umol_m2_sec, noon$flux_umol_m2_sec, paired = FALSE) # false bc the df's aren't equal length
t1
# 	Welch Two Sample t-test
# 
# data:  midnight$flux_umol_m2_sec and noon$flux_umol_m2_sec
# t = 2.5999, df = 877.99, p-value = 0.009482
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  0.1637867 1.1727094
# sample estimates:
# mean of x mean of y 
#  4.455344  3.787096 

# the two groups are statistically different from each other, ~425-460 observations each for midnight and noon; midnight significantly higher than noon, on average.  Not bad! also no random effect tho, need to keep that in mind; might be a good idea to separate by bot with a random effect.

# then, I want to compare all daytime data (8am-4pm = day, 8pm-4am = night).
day <- subset(fluxdat_final, hour==8|hour==9|hour==10|hour==11|hour==12|hour==13|hour==14|hour==15|hour==16) #3954 entries
# add col with 'day' marker
day$day_night <- "day"
night <- subset(fluxdat_final, hour==20|hour==21|hour==22|hour==23|hour==0|hour==1|hour==2|hour==3|hour==4) #3753 entries
# add col with 'night' marker
night$day_night <- "night"
t2 <- t.test(night$flux_umol_m2_sec, day$flux_umol_m2_sec, paired = FALSE)
t2
# t = 7.5646, df = 7692.4, p-value = 4.341e-14
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  0.4994595 0.8488599
# sample estimates:
#  mean of night  mean of day
#  4.409714       3.735554 

# night is also significantly higher flux than daytime hours
```

```{r mixedmodel}
# I also want to compare night/day and midnight/noon using a generalized linear mixed effects model that allows for the use of random effects (botid):
# is it temporally correlated?  is it repeated measures?  homogenous/constant variance, normality?
# boxplot of the data:
fluxdat_final %>% 
  subset(hour == 0 | hour == 12) %>% 
  ggplot(aes(x=as.factor(hour), y=flux_umol_m2_sec, fill=as.factor(hour)))+
  geom_violin()+
  geom_jitter(color="black", size=0.5, alpha=0.25)+
  theme_classic()+theme(legend.position = "none")+
  ggtitle("difference in soil flux between midnight (0) and noon (12)")

# okay now for day/night:
fluxdat_daynight <- rbind(night, day)
fluxdat_daynight %>% 
  ggplot(aes(x= day_night, y=flux_umol_m2_sec, fill=day_night))+
  geom_boxplot()+
  geom_jitter(color="black", size=0.5, alpha=0.25)+
  theme_classic()+theme(legend.position = "none")+
  ggtitle("difference in soil flux between day and night")

# definitely a visual difference in both cases. linear model of the day/night data 
lm1 <- lm(flux_umol_m2_sec~day_night+day_night*BOTID, 
          data=fluxdat_daynight)
plot(lm1) # not normal at all
summary(lm1)

lmer1 <- lmerTest::lmer(flux_umol_m2_sec~day_night + (1|BOTID), data = fluxdat_daynight)
plot(lmer1)
simulateResiduals(lmer1, plot = TRUE) # extremely high deviation on QQ plot, shows within-group heterogeneity too

# for the sake of simplicity going to go with the t-test for the methods paper; not looking to get super involved in the biological aspect of this analysis yet.

# visualize:
bxp <- ggboxplot(
  fluxdat_daynight, x = "day_night", y = "flux_umol_m2_sec",
  color = "treatment", palette = "jco"
  )
bxp

```

```{r repeatedmeasuresanova}
# I can also try two-way repeated measures anova, with an error term of BOTID.

# data is already in longitudinal form with the night/day operator in a column as well as BOTID
# https://www.r-bloggers.com/2015/08/two-way-anova-with-repeated-measures/
# first, find mean value for each BOTID per time of day; then, instead of one row per BOTID per observation, we'll have one row per BOTID per condition (day vs. night).
daynight_mean <- aggregate(fluxdat_daynight$flux_umol_m2_sec,
                           by = list(fluxdat_daynight$day_night, fluxdat_daynight$BOTID, fluxdat_daynight$treatment),
                           FUN = 'mean')
colnames(daynight_mean) <- c("day_night", "BOTID", "treatment", "mean_flux") #label the summarized columns

# build ANOVA:
# library(stats)

flux.aov <- with(daynight_mean,
                 aov(mean_flux ~ day_night + Error(BOTID/day_night)))
#model comparing flux by time of day, including an error term of BOTID for each condition (time of day) to look at how flux changes as a function of time of day, accounting for variation between individual BOTIDs. I tried this ANOVA with an interaction between day and night and treatment; that interaction was not significant so below summary is of JUST the mean_flux ~ day_night and an error term for BOTID

summary(flux.aov)

# Error: BOTID
#           Df Sum Sq Mean Sq F value Pr(>F)
# Residuals 11  25.14   2.285               
# 
# Error: BOTID:day_night
#           Df Sum Sq Mean Sq F value  Pr(>F)   
# day_night  1  3.475   3.475    19.4 0.00105 **
# Residuals 11  1.970   0.179     

bxp <- ggboxplot(
  daynight_mean, x = "day_night", y = "mean_flux",
  color = "treatment", palette = "jco"
  )
bxp

```
t-test of location:
```{r}
UT <- fluxdat_final %>% 
  filter(location == "Under Tree")
OS <- fluxdat_final %>% 
  filter(location == "Open Soil")

t3 <- t.test(UT$flux_umol_m2_sec, OS$flux_umol_m2_sec, paired = FALSE)
t3
# mean UT:  mean OS:
# 3.686133  4.311908
# p-value < 2.2e-16***

UT %>% group_by(day_night_ID) %>% summarize(mean = mean(flux_umol_m2_sec))
# day	  4.067619			
# night	3.769976
UT %>% t.test(flux_umol_m2_sec ~ day_night_ID, data = ., paired = FALSE)
# day: 4.067619            
# night: 3.769976 
# p=0.266

OS %>% group_by(day_night_ID) %>% summarize(mean = mean(flux_umol_m2_sec))
# day	  3.913344			
# night	4.946090	
OS %>% t.test(flux_umol_m2_sec ~ day_night_ID, data = ., paired = FALSE)
# day: 3.913344            
# night: 4.946090 
# p=0.004***

```
t-test of treatment:
```{r}
mwc <- fluxdat_final %>% 
  filter(treatment == "MWC")
o <- fluxdat_final %>% 
  filter(treatment == "O")

t4 <- t.test(mwc$flux_umol_m2_sec, o$flux_umol_m2_sec, paired = FALSE)
t4
# mean MWC:  mean O:
# 4.608967  3.401198 
# p-value < 2.2e-16***

mwc %>% group_by(day_night_ID) %>% summarize(mean = mean(flux_umol_m2_sec))
# day	  4.653717			
# night	4.959643
mwc %>% t.test(flux_umol_m2_sec ~ day_night_ID, data = ., paired = FALSE)
# day: 4.653717            
# night: 4.959643 
# p=0.408

o %>% group_by(day_night_ID) %>% summarize(mean = mean(flux_umol_m2_sec))
# day	  3.102450			
# night	3.861013	
o %>% t.test(flux_umol_m2_sec ~ day_night_ID, data = ., paired = FALSE)
# day: 3.102450            
# night: 3.861013 
# p=0.001

```

# temperature vs. flux relationship, day and night
This chunk explores (visually) the relationships between temp and flux for daytime and nighttime data
```{r tempflux}
# plot scatterplot of data with regressions between temp and flux, day and night hours:
fluxdat_daynight %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=avg_temp_degC, color=day_night))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, aes(fill=day_night), fullrange=TRUE)+
  theme_classic()

# midnight vs. noon:
fluxdat_final %>% 
  subset(hour == 0 | hour == 12) %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=avg_temp_degC, color=as.factor(hour)))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, aes(fill=as.factor(hour)), fullrange=TRUE)+
  theme_classic()

# with humidity instead:
fluxdat_daynight %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=avg_rel_humidity, color=day_night))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, aes(fill=day_night), fullrange=TRUE)+
  theme_classic()

# and pressure:
fluxdat_daynight %>% 
  ggplot(aes(x=flux_umol_m2_sec, y=avg_pressure_hPa, color=day_night))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, aes(fill=day_night), fullrange=TRUE)+
  theme_classic()
  
```


# Beta exploration
This chunk explores the beta values, generated by the linear and quadratic regressions run on each flux observation's raw data, including things like proportions of fluxes that were calculated with L vs Q betas, distributions of beta values, etc.
```{r betas} 
# distribution plots
fluxdat_final %>% 
  ggplot(aes(x=`X1st_order_beta_0`, fill=treatment, alpha=0.25))+
  geom_density()+
  theme_classic()

fluxdat_final %>% 
  ggplot(aes(x=`X2nd_order_beta_0`, fill=treatment, alpha=0.25))+
  geom_density()+
  theme_classic()

fluxdat_final %>% 
  ggplot(aes(x=regr, y=flux_umol_m2_sec))+
  geom_boxplot()+
  theme_classic() # plot exploring the distribution of flux values according to whether it was calculated with a linear or quadratic regression

fluxdat_final %>% 
  ggplot(aes(x=`X1st_order_beta_0`, fill=regr, alpha=0.25))+
  geom_density()+
  theme_classic()+
  ylim(c(0, 600000000))

fluxdat_final %>% 
  ggplot(aes(x=`X2nd_order_beta_0`, fill=regr, alpha=0.25))+
  geom_density()+
  theme_classic()+
  ylim(c(0, 600000000))

#######
# plot both on same plot; need long form data for that
library(reshape2)
fluxdat_beta_long <- melt(fluxdat_final,
                          id.vars = c("BOTID", "hour"),
                          measure.vars = c("X1st_order_beta_0",
                                           "X2nd_order_beta_0"),
                          variable.name = "beta_type",
                          value.name = "beta_value")
# plot:
ggplot(fluxdat_beta_long, aes(x=beta_value, group=beta_type, fill=beta_type))+
  geom_density(alpha=0.25)+
  facet_wrap(~BOTID)+
  theme_classic()

# do the same with day/night data:
fluxdat_beta_long_daynight <- melt(fluxdat_daynight,
                          id.vars = c("BOTID", "day_night"),
                          measure.vars = c("X1st_order_beta_0",
                                           "X2nd_order_beta_0"),
                          variable.name = "beta_type",
                          value.name = "beta_value")
# plot:
ggplot(fluxdat_beta_long_daynight, aes(x=beta_value, group=beta_type, fill=beta_type))+
  geom_density(alpha=0.25)+
  facet_wrap(~day_night)+
  theme_classic()

########
# also plot r-squareds against betas:
fluxdat_final %>% 
  ggplot(aes(x=`X1st_order_beta_0`, y=`X1st_order_r_sq`))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, fullrange=TRUE)+
  facet_wrap(~BOTID)+
  theme_classic()

fluxdat_final %>% 
  ggplot(aes(x=`X2nd_order_beta_0`, y=`X2nd_order_r_sq`))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm, fullrange=TRUE)+
  facet_wrap(~BOTID)+
  theme_classic()
```
```{r}
# count number of observations in cleaned dataset that are quadratic, and have a positive beta_2 (aka the slope increases over time):
fluxdat_final %>% 
  filter(regr == "Q") %>% 
  summarise(count = sum(X2nd_order_beta_1 > 0))
# 3 only three observations from the final dataset had a positive second beta (aka increasing slope over time)
```


# this chunk explores the R2s: distribution, which of the methods (Q vs. L) of calculating flux correspond to day vs. night the most often, what the R2s are for those curves, etc. etc. etc..
```{r R2sexploration}

# what I want to do is plot just the R2s for each method of calculation (e.g. 1st order R2s for L, 2nd order R2s for Q) and see the distributions for each.
fluxdat_final %>% 
  filter(regr == "Q") %>%  #filter operates on rows, select on cols
  ggplot(aes(x=`X2nd_order_r_sq`))+
  geom_density(alpha=0.25)+
  facet_wrap(~BOTID)+
  theme_classic()

# with day/night data:
fluxdat_daynight %>% 
  filter(regr == "Q") %>%  #filter operates on rows, select on cols
  ggplot(aes(x=`X2nd_order_r_sq`))+
  geom_density(alpha=0.25)+
  facet_wrap(~day_night)+
  theme_classic()

# basic exploration of R2s
mean(fluxdat_final$R2) #0.6466187
median(fluxdat_final$R2) #0.883
min(fluxdat_final$R2) #-24.182
max(fluxdat_final$R2) #0.999

# some more advanced exploration:
sum(fluxdat_final$R2 < 0) # 825 of dataset's final 'negative' R2's are less than zero
sum(fluxdat_final$R2 < -1) # 291 are less than -1
# percentage that are negative?
((sum(fluxdat_final$R2 < 0))/(sum(fluxdat_final$R2)))*100 # 12.62% of the R2s are negative
((sum(fluxdat_final$R2 > -0.25 & fluxdat_final$R2 < 0.25))/(sum(fluxdat_final$R2)))*100 # 9.75% of the negative R2s are sort of clustered around zero (-0.25 to 0.25)
((sum(fluxdat_final$R2 > 0.90))/(sum(fluxdat_final$R2)))*100 # 71.3% of the R2s are greater than 0.9, Carbone et al. threshold
cor(fluxdat_final$flux_umol_m2_sec, fluxdat_final$R2) # 23.9% correlation between R2 and flux estimate


```

Baaaasic descriptive stats of fluxbot data:
```{r descriptive_stats}
# mean of fluxbot data, all
mean(fluxdat_final$flux_umol_m2_sec) #4.050193

# median:
median(fluxdat_final$flux_umol_m2_sec) #3.144

# how many clustered around zero?
sum(fluxdat_final$flux_umol_m2_sec > -1 & fluxdat_final$flux_umol_m2_sec < 1)
# 1510, about 15%
```

Random plots for exploration:

Below, find a figure that plots R2 value (x axis) against flux, demonstrating that R2 increases as fluxes move away from zero.
```{r random}
fluxdat_final %>% 
  ggplot(aes(x=R2, y=flux_umol_m2_sec))+
  geom_point()+
  theme_classic()+
  labs(x = expression(paste("R"^2~"of regression used to calculate flux")),
       y = expression(paste("flux "~mu*"mol"~"/m"^2)*"/sec"))

```

