---
title: "rawdat_variability"
author: "Elizabeth Forbes"
date: "5/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries:
library(here)
library(tidyverse)
library(lubridate)
library(lme4)
library(nlme)
library(DHARMa)
library(lmerTest)
library(forecast)
library(TSdist)
library(ggseas)
library(changepoint)
library(Rbeast)
library(car)
```

# Raw data variability: did we observe increasing variability in data collection over time across the fluxbot array?
- upload raw data
- statistically analyze variance over time
- compare across fluxbots to see if any fluxbot(s) is driving an increase in variance, if this increase exists
```{r data}
# this chunk loads/tidies the humidity-adjusted (only) dataset
# import each file and then merge
filedir <- here::here("data", "raw accumulation data") # use "here" package/function to identify a sub-folder within the project's working directory
                      
# view files:
filedir %>% list.files() # these files are the 20s rolling average, humidity-corrected flux calculations

# save file names with extension .csv:
csv_filenames <- filedir %>% 
  list.files() %>% 
  .[str_detect(.,".csv")]

# load the files into the global environment of the project:
csv_filenames %>%
  purrr::map(function(file_name){ # iterate through each file name
    assign(x = str_replace(str_remove(file_name, ".csv"), " ", "_"), # Remove file extension ".csv", replace the space in the file names with "_"
           value = read_csv(file.path(filedir, file_name)),
           envir = .GlobalEnv)
  })

# add dataframe (and thus fluxbot) ID to data, bind into one
new_rawdat <- bind_rows(lst(NO_OS1, NO_OS2, NO_OS3, NO_UT1, NO_UT2, NO_UT3,
                            NMWC_OS1, NMWC_OS2, NMWC_OS3, NMWC_UT1, NMWC_UT2, NMWC_UT3),
                        .id = "botID")

# eliminate data from before August 23rd
new_rawdat <- new_rawdat %>% 
  filter(Timestamp > "2019-08-22 23:00:00")

# order the data
new_rawdat <- arrange(new_rawdat, Timestamp)

# add day, hour, month columns
new_rawdat <- new_rawdat %>% 
  mutate(hour = format(ymd_hms(as.character(new_rawdat$Timestamp)), '%H')) %>% 
  mutate(day = format(ymd_hms(as.character(new_rawdat$Timestamp)), '%d')) %>% 
  mutate(month = format(ymd_hms(as.character(new_rawdat$Timestamp)), '%m')) %>% 
  mutate(year = format(ymd_hms(as.character(new_rawdat$Timestamp)), '%y'))

# make exclusion df of just NO OS 1 data occurring after month == 9, i.e. October data:
exclude <- new_rawdat %>% 
  filter(botID == "NO_OS1") %>%
  filter(month == 10)

# use anti_join to remove those rows in 'exclude' from 'new_rawdat' (dplyr):
new_rawdat <- anti_join(new_rawdat, exclude)

# remove weird years
exclude2 <- new_rawdat %>% 
  filter(year != 19)
new_rawdat <- anti_join(new_rawdat, exclude2)

```

Calculate variance each day over time:

NOTE: this function may work, but the dataset is just too big.  It never finished running over the course of an hour or so.
```{r variance}
# https://stackoverflow.com/questions/61836898/variance-over-time-with-two-dataframes-in-r

# function to calculate cumulative variance, daily variance
# cumvar <- function(x)  {
#   sapply(seq_along(x), function(i) var(x[1:i]))
# }
# 
# test <- new_rawdat %>%
#   group_by(botID) %>%
#   mutate(cvar = cumvar(Filter.CO2.PPM)) %>%
#   group_by(botID, month) %>%
#   summarise(variance_month_x = var(Filter.CO2.PPM),
#             variance_cumulative = last(cvar)) %>%
#   complete(botID, month) %>%
#   fill(variance_cumulative, .direction = "down")

```

```{r changepoint}
# new_NO_OS1 <- arrange(NO_OS1, Timestamp)
# change_noos1 <- cpt.meanvar(new_NO_OS1$Filter.CO2.PPM)
# plot(change_noos1)
# 
# change_var <- cpt.var(new_rawdat$Filter.CO2.PPM, test.stat = "Normal",
#                       method = "PELT")
# plot(change_var)
# i have no idea how to interpret this
```

just going to do some classic f-tests here (more accurately Levene tests bc I have more than two )
```{r f_test}
# f test, with multiple independent variables (month and botID):
f1 <- leveneTest(Filter.CO2.PPM~month*botID, data = new_rawdat)
f1

f2 <- leveneTest(Filter.CO2.PPM~month, data = new_rawdat)
f2
# Levene's Test for Homogeneity of Variance (center = median)
#            Df F value    Pr(>F)    
# group       2   36270 < 2.2e-16 ***
#       3535048 
# variance apparently differs significantly across months.  hmm!
boxplot(Filter.CO2.PPM ~ month,
        data = new_rawdat,
        xlab = "month",
        ylab = "raw CO2 concentration, ppm",
        col = "steelblue",
        border = "black")

f3 <- leveneTest(Filter.CO2.PPM ~ botID, data = new_rawdat)
f3
# Levene's Test for Homogeneity of Variance (center = median)
#            Df F value    Pr(>F)    
# group      11  484778 < 2.2e-16 ***
#       3535039 

# variance also differs across botID.
boxplot(Filter.CO2.PPM ~ botID,
        data = new_rawdat,
        xlab = "fluxbot ID",
        las = 1.25,
        ylab = "raw CO2 concentration, ppm",
        col = "tomato1",
        border = "black")

# violin plot of the same data: first bot, then month
new_rawdat %>% 
  ggplot(aes(x=botID, y=Filter.CO2.PPM, fill = botID))+
  geom_violin()

month.labs <- c("Sept 2019", "Oct 2019", "Nov 2019")
names(month.labs) <- c("08", "09", "10")

new_rawdat %>% 
  ggplot(aes(x=month, y=Raw.CO2.PPM, fill = month))+
  geom_violin()+
  labs(y="Raw carbon dioxide concentration (ppm)",
       x="fluxbot ID")+
  theme_classic()+
  scale_fill_discrete()+
  theme(legend.position = "none")+
  # facet_wrap(~month, labeller = labeller(month=month.labs))+
  theme(axis.text.x = element_text(angle = 35, hjust=1))
```
violin plot of raw CO2 across months:
```{r}
new_rawdat %>% 
  ggplot(aes(x=month, y=Raw.CO2.PPM))+
  geom_boxplot(outlier.alpha = 0.25, fill = "steelblue")+
  # geom_violin()+
  labs(y="Raw carbon dioxide concentration (ppm)",
       x="")+
  theme_classic()+
  scale_x_discrete(labels=c("Sept. 2019", "Oct. 2019", "Nov. 2019"))+
  theme(legend.position = "none")
  # facet_wrap(~month, labeller = labeller(month=month.labs))+
  # theme(axis.text.x = element_text(angle = 35, hjust=1))

```

